PyQBench: a Python library for benchmarking gate-based quantum computers

### 2023-01-30 txt/manuscript.tex
pyqbench as library done

\documentclass[preprint,12pt, a4paper]{elsarticle}

@@

\usepackage{lineno}

@@

\restylefloat{table}

\newcommand{\eg}{{\emph{e.g.\/}}}
\newcommand{\ie}{{\emph{i.e.\/}}}
\newcommand{\ket}[1]{\ensuremath{|#1\rangle}}
\newcommand{\bra}[1]{\ensuremath{\langle#1|}}
\newcommand{\ketbra}[2]{\ensuremath{\ket{#1}\bra{#2}}}
\newcommand{\proj}[1]{\ensuremath{\ketbra{#1}{#1}}}
\newcommand{\braket}[2]{\ensuremath{\langle{#1}|{#2}\rangle}}
\newcommand{\floor}[1]{\ensuremath{\lfloor #1 \rfloor}}
\newcommand{\complexity}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\new}[1]{ \textcolor{red}{#1} }
\newcommand{\1}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\Id}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\connected}{\sim}
\newcommand{\SPAN}{\mathrm{span}}
\newcommand{\Lrm}{\ensuremath{\mathrm{L}}}
\newcommand{\Urm}{\ensuremath{\mathrm{U}}}
\newcommand{\ee}{\ensuremath{\mathrm{e}}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\ii}{\ensuremath{\mathrm{i}}}
\newcommand{\EE}{\mathcal{E}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\MM}{\mathcal{M}}
\newcommand{\NN}{\mathcal{N}}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\TT}{\mathcal{T}}
\newcommand{\PP}{\mathcal{P}}
\renewcommand{\SS}{\mathcal{S}}
\newcommand{\UU}{\mathcal{U}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\DU}{\mathcal{DU}}
\newcommand{\NOT}{\sigma_x}
\newcommand{\idop}[1][\XX]{\ensuremath{\1_{#1}}}
\newcommand{\diaguni}{\ensuremath{\mathcal{DU}}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{tr}}
\newcommand{\textapprox}{\raisebox{0.5ex}{\texttildelow}}
\journal{SoftwareX}

@@

\begin{table}[H]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Please fill in this 
column} \\
\hline
C1 & Current code version & 0.1.0 \\
\hline
C2 & Permanent link to code/repository used for this code version & \url{https://github.com/iitis/PyQBench} \\
\hline
C3 & Code Ocean compute capsule & ?????????\\
\hline
C4 & Legal Code License & Apache License 2.0\\
\hline
C5 & Code versioning system used & git \\
\hline
C6 & Software code languages, tools, and services used & Python, Qiskit, AWS Braket \\
\hline
C7 & Compilation requirements, operating environments \& dependencies & 
\texttt{Python >= 3.8}\newline
\texttt{amazon-braket-sdk \textapprox= 1.33.0}\newline
\texttt{numpy \textapprox= 1.14.5}\newline
\texttt{scipy \textapprox= 1.7.3}\newline
\texttt{pandas \textapprox= 1.5.0}\newline
\texttt{pydantic \textapprox= 1.9.0}\newline
\texttt{qiskit \textapprox= 0.39.0}\newline
\texttt{mthree \textapprox= 1.1.0}\newline
\texttt{tqdm \textapprox= 4.64.1}\newline
\texttt{pyyaml \textapprox= 6.0}\\
\hline
C8 & If available Link to developer documentation/manual & ?????\\
\hline
C9 & Support email for questions & \url{dexter2206@gmail.com}\\
\hline
\end{tabular}
\caption{Code metadata (mandatory)}
\label{} 
\end{table}

@@

\section{Software description}
\label{}
\todo[inline]{Describe the software in as much as is necessary to establish a 
vocabulary needed to explain its impact. }

We will first give a general overview of the structure of the code in 
Section~\ref{sec:sortware-architecture} and then provide additional details on 
the functionality of benchmark's schemes  in 
Section~\ref{sec:sortware-functionalities}.

\subsection{Software Architecture}\label{sec:sortware-architecture}
\label{}
\todo[inline]{
	Give a short overview of the overall software architecture; provide a 
	pictorial component overview or similar (if possible). If necessary provide 
	implementation details.}

\subsection{Software Functionalities}\label{sec:sortware-functionalities}
\todo[inline]{Present the major functionalities of the software.}

\subsubsection{Model specification}
The scenario assumes that Alice and Bob have access to an unknown measurement
device, a black box. Alice has hidden one of the two measurements, either $\PP$
or $\mathcal{Q}$, in the black box. Bob only has the information that the black box
performs one of these measurements. They want to determine an upper bound on the
probability of correct discrimination between $\PP$ and $\mathcal{Q}$. Moreover, they also want to
construct an optimal discrimination strategy to achieve the highest probability. Theoretical results were characterized in~\cite{puchala2018strategies}.

In this work, we will consider the problem of discrimination between von Neumann 
measurements. One of them on the computational basis, the second one on basis 
$U$, where $U$ is a unitary matrix. 
The scheme of discrimination can be divided into the following steps, which are 
illustrated in Fig.~\ref{fig:theoretical_scheme}.

@@

%\begin{figure}[h!]
%	\centering 
%	\includegraphics[width=0.9\linewidth]{gates1.pdf} 
%	\caption{A schematic representation of the setup for distinguishing
%		measurements by using entanglement. }
%	\label{fig:scheme}
%\end{figure}

\begin{enumerate}
	\item Alice and Bob share an entangled state $\ket{\psi_{0}}$ (discriminator) on systems 
	$A$ and $B$.
	\item Alice performs one of two known measurements (either $\PP_{\Id}$ or 
	$\PP_{U}$) on part $A$ of the input state  $\ket{\psi_{0}}$ and measures the system $A$.
	\item Alice performs the conditional binary measurement $\PP_{V_i}$ on the part 
	$B$ based on the output label $i$. 
	\item  Bob measures the system $B$ of the input state  $\ket{\psi_{0}}$ and 
	makes a decision based on the received label $j$. If $j=0$, then he decides 
	that the the black box contains $\PP_U$. Otherwise, he decides that the 
	black box contains $\PP_{\Id}$. 
\end{enumerate}

\subsubsection{Quantum architecture's limits and practical solutions}

Let us take a closer look at the setup for distinguishing von Neumann 
measurements (see Fig.~\ref{fig:theoretical_scheme}). One of the components of 
this scheme is to perform a conditional binary measurement. 
Unfortunately, current NISQ devices do not allow for performing conditional 
binary measurements.
The only possible measurement performed on quantum architectures is the 
measurement in the computational basis. 
We  will present two possible methods providing a clever way to cross this 
obstacle. These will be two schemes of practical realization. 
The first method consists in using postselection while the second method 
consists in applying a controlled unitary. These methods are described below.

\begin{scheme}(Postselection)

%\begin{figure}[h!]
%	\includegraphics[scale=1.5]{onequbit.pdf}
%	\caption{A schematic representation of the setup for distinguishing
%		measurements using postselection.}
%	\label{postselection}
%\end{figure}

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{0}}$ (discriminator) on systems 
$A$ and $B$.
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or 
$\Phi_{U^\dagger}$,  on part $A$ of the input state  $\ket{\psi_{0}}$.
\item Alice measures part $A$ of the input state  $\ket{\psi_{0}}$ and 
receives an output label $i$.
\item  
Based on Holevo--Helstrom theorem, Bob performs a conditional binary 
measurement	$\PP_{V_j}$ and implements the unitary channel 	$\Phi_{V_j^\dagger}$ on part $B$.
\item To calculate the probability of correct discrimination, they include just those cases for which $i = j$.
\item Bob measures part $B$ of the input state  $\ket{\psi_{0}}$ and makes a
decision based on received label $j$. If $i=j=0$, then he decides that the 
black box contains $\PP_U$. Otherwise, he decides that the black box contains
$\PP_{\Id}$.
\end{enumerate}

The schematic representation of this setup is depicted in 
Fig.~\ref{fig:postsellection}.

\begin{figure}[h!]
	\centering 
	\includegraphics[scale=1.7]{pics/postselection}

	\caption{ A schematic representation of the setup for distinguishing
		measurements using postselection. 
	}\label{fig:postsellection}
\end{figure}

Due to the fact that the measurement result is nondeterministic, we are forced to consider the following circuits
in the postselection procedure
\begin{equation}
\begin{split}
(\Phi_{U^\dagger}, \Phi_{V_k^\dagger}, i, j) \\ 
(\Phi_{\Id}, \Phi_{V_k^\dagger}, i, j)
\end{split}
\end{equation}
where $k=\{0,1\}, i=\{0,1\}$ and $ j=\{0,1\}$. 
The optimal probability of correct discrimination between $\PP_{U} $ and $\PP_\Id$ is as follows.
		\begin{equation}
		p = \frac{\#(\Phi_{U^\dagger}, \Phi_{V_i^\dagger},i,0) + \#(\Phi_\Id, \Phi_{V_i^\dagger},i,1)}{\#(\Phi_{U^\dagger}, \Phi_{V_i^\dagger},i,j) + \#(\Phi_\Id, \Phi_{V_i^\dagger},i,j)}, 
		\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$.

@@

\begin{scheme}(By using controlled unitary)

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{0}}$ on 
systems $A$ and $B$.
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or
$\Phi_{U^\dagger}$,  on part $A$ of the input state  $\ket{\psi_{0}}$.
\item Alice and Bob implement  direct sum of unitary channels $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger}$ obtained from 
Holevo--Helstrom theorem.	
\item Alice and Bob prepare the measurement $\Delta$ in computational basis on 
their systems.
\item Bob makes a decision based on received label $j$. If $j=0$, then he
decides that the the black box contains $\PP_U$. Otherwise, he decides that the
black box contains $\PP_{\Id}$.
\end{enumerate}

@@

Here, we consider the following circuits
by using controlled unitary
\begin{equation}
\begin{split}
(\Phi_{U^\dagger}, \Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger}, i, j) \\ 
(\Phi_{\Id},  \Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger}, i, j)
\end{split}
\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$. 
Then, the optimal probability of correct discrimination between $\PP_{U} $ and $\PP_\Id$ is as follows.
\begin{equation}
p = \frac{\#(\Phi_{U^\dagger},\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger},i,0) + \#(\Phi_\Id,\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger},i,1) }{\#(\Phi_{U^\dagger},\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger},i,j) + \#(\Phi_\Id,\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger},i,j)}
\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$. 
\end{scheme}

\subsubsection{Mathematical preliminaries}

Let $M_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity, square matrices will be denoted by
$M_d$.  
%The subset of $M_d$ consisting of Hermitian matrices of dimension $d$
%will  be  denoted  by $\HH_d$,  while  the  set  of  positive semidefinite
%matrices of dimension $d$ by $\HH_d^+$. 
The set of quantum states, that is
positive semidefinite operators having trace equal to one, will be denoted
$\Omega_d$. 
 The subset of $M_d$ consisting of unitary matrices will be denoted
by $\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$. 
	An operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a stochastic matrix if $a_{i,j} \ge 0$   and $\sum_{i} a_{i,j} = 1$, whereas an operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a double stochastic matrix if $A$ is a stochastic matrix and $\sum_{j} a_{i,j} = 1$.

 We will also need a linear mapping transforming $M_{d_1}$ into
$M_{d_2}$, which will be denoted 
\begin{equation}
\Phi: M_{d_1 } \rightarrow M_{d_2}.
\end{equation} 
There
exists a bijection between the set of linear mappings $\Phi$ and the set of matrices $M_{d_1d_2}$,  known as the Choi-Jamio{\l}kowski isomorphism. 
For a given linear mapping $\Phi$ the corresponding Choi operator $\mathcal{J}(\Phi)$ is explicitly written as 
\begin{equation}
\mathcal{J}(\Phi) \coloneqq \sum_{i,j=0}^{d- 1} \Phi(\ketbra{i}{j}) \otimes \ketbra{i}{j}. \end{equation}
We introduce a special subset of all mappings $\Phi$, called quantum channels, which are completely positive
and trace preserving (CPTP).
In this work we will consider a special class of quantum channels, called unitary channels.  A 
channel
$\Phi_{U}$ is said to be a unitary channel if it has the following form $\Phi_U(\cdot) = U \cdot U^\dagger$ for  $U \in 
\UU_d$.

Finally, we define a quantum
measurement, that is a positive operator valued measure (POVM) denoted by $\PP, \mathcal{Q}$ etc. This is a 
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
effects, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$. If
all effects are rank-one projection operators, then such a measurement is
called a von Neumann measurement. 
Every von Neumann measurement can be
parameterized by a unitary matrix $U$ with effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
where $\ket{u_i}$ is the $i$-th column of the unitary matrix $U$.  Hence we will use the notation $\PP_{U}$.  The action of
quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
seen as  a measure-and-prepare quantum channel as follows 
$
\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
$ Moreover, observe that each von Neumann measuement $\PP_{U}$ can be written as compostion of a unitary channel $\Phi_{U^\dagger}$ and the maximally dephasing channel $\Delta$, that means $\PP_{U} = \Delta \circ \Phi_{U^\dagger}$.

\subsubsection{Diamond norm and the task of discrimination between quantum operators}
This section defines a special norm on the space of linear mappings, known as 
the diamond norm, and establishes some of its properties. 
Next, we will explain the precise connection between the diamond
norm and the task of quantum operation discrimination.

>>>>

\documentclass[preprint,12pt, a4paper, dvipsnames]{elsarticle}

@@

\usepackage{lineno}
\usepackage{listings}

@@

\usepackage{xcolor}

\usepackage{caption}
\usepackage{subcaption}

\definecolor{codegreen}{rgb}{0.25,0.93,0.31}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

%
%\lstdefinestyle{mystyle}{
%	backgroundcolor=\color{backcolour},   
%	commentstyle=\color{codegreen},
%	keywordstyle=\color{magenta},
%	numberstyle=\tiny\color{codegray},
%	stringstyle=\color{codepurple},
%	basicstyle=\ttfamily\footnotesize,
%	breakatwhitespace=false,         
%	breaklines=true,                 
%	captionpos=b,                    
%	keepspaces=true,                 
%	numbers=left,                    
%	numbersep=5pt,                  
%	showspaces=false,                
%	showstringspaces=false,
%	showtabs=false,                  
%	tabsize=2
%}
%
%\lstset{style=mystyle} 
\lstset{frame=tb,
	language=Python,
	aboveskip=3mm,
	belowskip=3mm,
	showstringspaces=false,
	columns=flexible,
	basicstyle={\small\ttfamily},
	numberstyle=\color{blue},
	keywordstyle=\color{blue},
	commentstyle=\color{PineGreen},
	stringstyle=\color{ForestGreen},
	breaklines=true,
	breakatwhitespace=true,
	tabsize=3,
	escapeinside={@}{@},
	literate=
	{0}{{{\color{red}0}}}1
	{1}{{{\color{red}1}}}1
	{2}{{{\color{red}2}}}1
	{3}{{{\color{red}3}}}1
	{4}{{{\color{red}4}}}1
	{5}{{{\color{red}5}}}1
	{6}{{{\color{red}6}}}1
	{7}{{{\color{red}7}}}1
	{8}{{{\color{red}8}}}1
	{9}{{{\color{red}9}}}1
    {v0}{{{\color{black}v0}}}1 %% we can not style only numeric literals
    {v1}{{{\color{black}v1}}}1
}
\restylefloat{table}

\newcommand{\eg}{{\emph{e.g.\/}}}
\newcommand{\ie}{{\emph{i.e.\/}}}
\newcommand{\ket}[1]{\ensuremath{|#1\rangle}}
\newcommand{\bra}[1]{\ensuremath{\langle#1|}}
\newcommand{\ketbra}[2]{\ensuremath{\ket{#1}\bra{#2}}}
\newcommand{\proj}[1]{\ensuremath{\ketbra{#1}{#1}}}
\newcommand{\braket}[2]{\ensuremath{\langle{#1}|{#2}\rangle}}
\newcommand{\floor}[1]{\ensuremath{\lfloor #1 \rfloor}}
\newcommand{\complexity}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\new}[1]{ \textcolor{red}{#1} }
\newcommand{\1}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\Id}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\connected}{\sim}
\newcommand{\SPAN}{\mathrm{span}}
\newcommand{\Lrm}{\ensuremath{\mathrm{L}}}
\newcommand{\Urm}{\ensuremath{\mathrm{U}}}
\newcommand{\ee}{\ensuremath{\mathrm{e}}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\ii}{\ensuremath{\mathrm{i}}}
\newcommand{\EE}{\mathcal{E}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\MM}{\mathcal{M}}
\newcommand{\NN}{\mathcal{N}}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\TT}{\mathcal{T}}
\newcommand{\PP}{\mathcal{P}}
\newcommand{\QQ}{\mathcal{Q}}
\renewcommand{\SS}{\mathcal{S}}
\newcommand{\UU}{\mathcal{U}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\DU}{\mathcal{DU}}
\newcommand{\NOT}{\sigma_x}
\newcommand{\idop}[1][\XX]{\ensuremath{\1_{#1}}}
\newcommand{\diaguni}{\ensuremath{\mathcal{DU}}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{tr}}
\newcommand{\textapprox}{\raisebox{0.5ex}{\texttildelow}}
\journal{SoftwareX}

@@

\begin{table}[H]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Please fill in this 
column} \\
\hline
C1 & Current code version & 0.1.0 \\
\hline
C2 & Permanent link to code/repository used for this code version & \url{https://github.com/iitis/PyQBench} \\
\hline
C3 & Code Ocean compute capsule & \todo[inline]{???}\\
\hline
C4 & Legal Code License & Apache License 2.0\\
\hline
C5 & Code versioning system used & git \\
\hline
C6 & Software code languages, tools, and services used & Python, Qiskit, AWS Braket \\
\hline
C7 & Compilation requirements, operating environments \& dependencies & 
\texttt{Python >= 3.8}\newline
\texttt{amazon-braket-sdk \textapprox= 1.33.0}\newline
\texttt{numpy \textapprox= 1.14.5}\newline
\texttt{scipy \textapprox= 1.7.3}\newline
\texttt{pandas \textapprox= 1.5.0}\newline
\texttt{pydantic \textapprox= 1.9.0}\newline
\texttt{qiskit \textapprox= 0.39.0}\newline
\texttt{mthree \textapprox= 1.1.0}\newline
\texttt{tqdm \textapprox= 4.64.1}\newline
\texttt{pyyaml \textapprox= 6.0} \todo[inline]{everything???}\\
\hline 
C8 & If available Link to developer documentation/manual & \todo[inline]{???}\\
\hline
C9 & Support email for questions & \url{dexter2206@gmail.com}\\
\hline
\end{tabular}
\caption{Code metadata (mandatory)}
\label{} 
\end{table}

@@

!!!!!!!

Suppose we have access to a device performing one of the predefined von Neumann
measurements, $\PP$ or $\QQ$. In PyQBench, we restrict ourselves to discriminating von Neumann measurements.
This is because, unlike other measurement types, they can be implemented on actual hardware. While the $\PP$ and $\QQ$ are known, it is not known which of them is
performed when the device executes. Based on the measurement outcome, you have to guess which
measurement was performed (but you can perform arbitrary unitary operations before and after the
measurement). What is the highest probability of making a correct guess? And what do you have to do
to achieve it? And, most importantly, why would we want to do this?

Suppose that you know a strategy that, for an ideal device, would yield a probability
$p_{\text{succ}}$ of successfully discriminating between two measurements. Will the probability be
the same on an actual physical device? For current, Noisy Intermediate Scale Quantum devices
(NISQs), the answer is: certainly not. However, the error rate that you make when guessing can be
used as a benchmarking metric. PyQBench helps you in organizing such discrimination experiments for
a single qubit system, executing them on real hardware or simulators, and
computing discrimination probabilities based on the measured bitstrings.

\section{Preliminaries and discrimination scheme approach}
This section is divided into two parts.
In Section~\ref{sec:maths}, we introduce the necessary notation and preliminaries needed to understand the concept of PyQBench. Next, in Section~\ref{sec:discrimination-scheme}, 
we will present a general overview of the discrimination scheme in theoretical terms. We will also mention NISQ devices' limitations and possible practical solutions in the implementation process.

\subsection{Mathematical preliminaries}\label{sec:maths}
Recall that a general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$.
In PyQBench, we are interested only in von Neumann measurements, i.e. measurements
for which all the effects are rank-one projectors. Every such measurement can be
parameterized by a unitary matrix $U$ which the effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
are created by taking $\ket{u_i}$ as  $i+1$-th column of the unitary matrix $U$.
We will denote von Neumann measurements described by the matrix $U$ by $\PP_{U}$.

Typically, NISQ devices can only perform measurements in computational $Z$-basis.
To perform an arbitrary von Neumann measurement $\PP_{U}$, one has to first apply $U^\dagger$
to the measured system and then follow with $Z$-basis measurement. Hence, the following two
circuits are equivalent.

\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/vonneuman}
	\caption{Decomposition of von Neumann measurement}
	\label{fig:vonnneuman}
\end{figure}

Note that the decomposition above is basically a change of basis in which the measurement
is performed.

\subsection{Discrimination scheme}\label{sec:discrimination-scheme}

Without loss of generality, we consider discrimination between a measurement in the computational
Z-basis ($\PP_\Id$), and an alternative measurement performed in the basis $U$
($\PP_U$)\footnote{Explaining why we can consider only Z-basis and alternative measurement is beyond
	the scope of this technical documentation. See \cite{puchala2018strategies} if you are interested in
	the explanation.}. In PyQBench, we operate only on two-level systems, but the discrimination scheme,
described in detail in \cite{puchala2018strategies}, makes no assumptions about the dimensionality.

In general, the discrimination scheme, presented in Fig.\ref{fig:theoretical_scheme}, requires a
second system of the same dimensionality as the measured one. First, the joint system is prepared in
some state $\ket{\psi_0}$. Then, the unknown measurement is performed on the first part of the
system. Based on its outcome $i$, another measurement $\mathcal{P}_{V_i}$ is performed to obtain an
outcome $j$. Finally, if $j=0$ we guess that the performed measurement is $\mathcal{P}_U$, otherwise
we guess that it was $\mathcal{P}_\Id$.

@@

\subsubsection{Limitations of NISQ devices and practical discrimination schemes}

Current NISQ devices are unable to perform conditional measurements, which is the biggest
obstacle to implementing our scheme on real hardware. Luckily, we can overcome it by
cleverly adjusting our scheme so that it only uses components available on current devices.
We have two possible choices here: using a postselection or a direct sum
$V_0^\dagger\oplus V_1^\dagger$.

\begin{scheme}(By using postselection)

	The first idea is very simple. Instead of performing a conditional measurement, for each
	measurement $\PP_U, \PP_\Id$ to be discriminated and each choice of $k \in \{0, 1\}$ we run circuit presented in Fig. \ref{fig:postselection}.

	\begin{figure}[h!]
		\centering
		\includegraphics[scale=1.7]{pics/postselection_no_channels}

		\caption{
			Postselection scheme of discrimination between von Neumann measurements $\PP_{U}$ and $\PP_\Id$.
		}\label{fig:postselection}
	\end{figure}

	But now our experiment does not match the previously described scheme, right?
	Yes, unless we discard all the outcomes for which $i\ne k$ (hence the name \emph{postselection},
	we are selecting valid outcomes after the experiment is performed).

	% \begin{scheme}(Postselection)

	% %\begin{figure}[h!]
	% %	\includegraphics[scale=1.5]{onequbit.pdf}
	% %	\caption{A schematic representation of the setup for distinguishing
	% %		measurements using postselection.}
	% %	\label{postselection}
	% %\end{figure}

	% \begin{enumerate}
	% \item We prepare  the discriminator $\ket{\psi_{0}}$ on bipartite system.
	% \item We perform one of two quantum channels, either $\Phi_{\Id}$ or
	% $\Phi_{U^\dagger}$,  on  the first part of the input state  $\ket{\psi_{0}}$, that means $\left( \Phi \otimes \Id\right)(\proj{\psi_0})$.
	% \item We measure the first part of the input state  $\ket{\psi_{0}}$ and
	% receive an output label $i$.
	% \item
	% Based on Holevo--Helstrom theorem, we perform a conditional binary
	% measurement	$\PP_{V_j}$ and hence, we implement the quantum channel $\Phi_{V_j^\dagger}$ on the second part.
	% \item To calculate the probability of correct discrimination, we include just those cases for which $i = j$.
	% \item We measure the second part of the input state  $\ket{\psi_{0}}$ and make a
	% decision based on received label $j$. If $i=j=0$, then we decide that
	% $\PP_U$ occurs.  If $i=j=1$, we decide that
	% $\PP_{\Id}$ occurs.
	% \end{enumerate}

	More precisely, our experiments can be grouped into classes identified by tuples of the form
	$(\mathcal{Q}, k, i, j)$, where $\mathcal{Q} \in \{\PP_U, \PP_\Id\}$ denotes the chosen measurement.
	We discard all the experiments for which $k \ne i$. Hence, the total number of valid experiments is

	\begin{equation}
	N_\text{total} = \#\{(\QQ, k, i, j): k = i \}. 
	\end{equation}

	We now need to count the experiments (among the valid ones) resulting in successful discrimination.
	If we have chosen $\PP_U$, then we guess correctly iff $j=0$. Similarly, for
	$P_\Id$, we guess correctly iff $j=1$. If we define
	\begin{eqnarray}
	N_{\PP_U} &= \#\{(\mathcal{Q}, k, i, j): \mathcal{Q} = \PP_U, k = i, j = 0\}, \\
	N_{\PP_\Id} &= \#\{(\mathcal{Q}, k, i, j): \mathcal{Q} = \PP_\Id, k = i, j = 1\},
	\end{eqnarray}
	then the empirical success probability can be computed as

	\begin{equation}
	p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{N_{\PP_U} + N_{\PP_\Id}}{N_{\text{total}}}.
	\end{equation}

@@

\begin{scheme}(By using direct sum)

	\begin{enumerate}
		\item We prepare the discriminator $\ket{\psi_{0}}$ on bipartite
		system.
		%\item We perform one of two quantum channels, either $\Phi_{\Id}$ or
		%$\Phi_{U^\dagger}$,  on  the first part of the input state  $\ket{\psi_{0}}$, that means $\left(\Phi \otimes \Id\right)(\proj{\psi_0})$.
		\item We apply one of unitary,   either $U^\dagger$ or
		$\Id$, on the first system.
		%\item Based on Holevo--Helstrom theorem, we perform a conditional binary measurement $\PP_{V_j}$, and hence, we implement the direct sum of quantum channels $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger}$. Observe that   $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger} = \proj{0} \otimes \Phi_{V_0^\dagger}  + \proj{1} \otimes \Phi_{V_1^\dagger}.  $
		\item We apply direct sum $V_0^\dagger \oplus V_1^\dagger$ on the whole systems.
		\item We measure both systems in computational basis $\Delta$.
		\item We make a decision based on the received label $j$ on the second system. If $j=0$, then we
		decide that $\PP_U$ occurs. Otherwise, we decide that $\PP_{\Id}$ occurs.
	\end{enumerate}

@@

	In this scheme, the experiment can be characterized by a pair $(\mathcal{Q}, i,j)$, where $\mathcal{Q} = \{ \PP_{U}, \PP_{\Id} \}$. The number of successful trials for $U$ and $\Id$, respectively, can be written  as
	\begin{eqnarray}
	N_{\PP_U} &= \#\{(\mathcal{Q},  i, j): \mathcal{Q} = \PP_U, j = 0\}, \\
	N_{\PP_\Id} &= \#\{(\mathcal{Q},  i, j): \mathcal{Q} = \PP_\Id, j = 1\}.
	\end{eqnarray}
	Then, the probability of correct discrimination between $\PP_{U} $ and $\PP_\Id$ is given by
	\begin{equation}
	p_{\text{succ}} = \frac{N_{\PP_{U}} + N_{\PP_{\Id}}}{N_{\text{total}}},
	\end{equation}
	where $N_{\text{total}}$ is the number of trials.
\end{scheme}

\subsubsection{Optimal discrimination scheme}

The celebrated result by Helstrom~\cite{helstrom1976quantum} gives the optimal  probability of correct discrimination between two quantum measurements, $\PP$  and $\mathcal{Q}$,
in terms of their distance with the use of the diamond norm
\begin{equation}
p_{\text{succ}}(\PP, \mathcal{Q}) =  \frac12 + \frac14 \| \PP - \mathcal{Q} \|_\diamond, 
\end{equation}
where 
\begin{equation}
\|\PP\|_\diamond = \max_{\| \ket{\psi}\|_1=1} \| \left(\PP \otimes \1\right) (\proj{\psi}) \|_1.
\end{equation}
The quantum state $\ket{\psi}$ which maximizes the diamond norm is called as discriminator.
Furthermore,  thanks to the proof of the Holevo-Helstrom theorem, it is possible to construct the necessary components  ($\ket{\psi_0}$, $V_0$, $V_1$) to create the optimal discrimination strategy.

 \section{Software description}
 \label{}
 This section is divided into two parts.
 In Section~\ref{sec:sortware-functionalities} we present the details of PyQBench functionality. Next, in Section~\ref{sec:sortware-architecture}, 
 we will give a general overview of the software architecture.

\subsection{Software Functionalities}\label{sec:sortware-functionalities}

\subsection{Software Architecture}\label{sec:sortware-architecture}
PyQBench can be used in two different ways:
\begin{itemize}
\item As a library. This is a more extensible, but also more laborious way of using PyQBench. Most importantly, this way allows for more advanced scripting defining your own measurements scheme.

\item As a CLI tool. This mode of operation trades extensibility for simplicity of usage, as it allows only for running Fourier-discrimination experiments defined with simple YAML input files.
\end{itemize}

\subsubsection{PyQBench as a library}\label{sec:pyqbench-as-library}
This guide introduces basic concepts needed for using PyQBench as a library from a Python script. In particular, we will cover the following topics:
\begin{enumerate}
\item Defining experiment using circuit components.
\item Assembling circuits needed for given benchmarking scheme.
\item Running experiment on simulator or hardware.
\item Obtaining empirical probability of successful discrimination between measurements.
\end{enumerate}

On the conceptual level, every discrimination experiment performed in PyQBench needs the following:
\begin{enumerate}
\item Discriminator, i.e. optimal initial state for the circuit.
\item Unitary $U^\dagger$ defining alternative measurement to be discriminated from the measurement in Z-basis.
\item $V_0^\dagger$ and $V_1^\dagger$, positive and negative parts of optimal measurement from Holevo-Helstrom theorem.
\end{enumerate}
Currently available quantum computers typically cannot start execution from an arbitrary state. Instead, we have to prepare it ourselves. Hence, to execute experiment using postselection scheme, we need an instruction taking to the desired discriminator
and instructions implementing $U^\dagger$, $V_0^\dagger$ and $V_1^\dagger$.

\subsubsection{Example of usage}
In our example, we will use $U =  H$ (the Hadamard gate). To keep us focused on the implementation in PyQBench and not delve too deep into mathematical explanation, we simply provide explicit formulas for discriminator $\ket{\Psi_0} $, $V_0$ and $V_1$, leaving the calculations to the interested reader.

The explicit formula for discriminator in our toy model reads:
\begin{equation}
\ket{\Psi_0} = \frac{1}{\sqrt{2}} (\ket{00} + \ket{11}), 
\end{equation}
with corresponding parts of optimal measurement being equal to
\begin{equation}
V_0 = 
\left(\begin{array}{cc} \alpha & -\beta\\ \beta & \alpha \end{array}\right), 
\end{equation}
and \begin{equation}
V_1 = 
\left(\begin{array}{cc} -\beta & \alpha \\ \alpha & \beta \end{array}\right), 
\end{equation} 
where \begin{equation}
\alpha = \frac{\sqrt{2 - \sqrt{2}}}{2} \cos\left( \frac{3}{8} \pi \right),
\end{equation}
\begin{equation}
\beta  = \frac{\sqrt{2  + \sqrt{2}}}{2} \sin\left( \frac{3}{8} \pi \right).
\end{equation}
